Titeln på avhandlingens språk

### 2023-02-03 publications/2022Ainur/ainur.tex
fixing formatting

\section{The Ainur Framework}\label{paper:olguinmunoz2022airnur:ainur}

\begin{figure}[t]
    \centering
    \includegraphics[width=\textwidth]{publications/2022Ainur/figures/overview}
    \caption{Layered structure of an Ainur experiment}\label{paper:olguinmunoz2022airnur:fig:overview}
\end{figure}

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{publications/2022Ainur/figures/flow2}
    \caption{Lifecycle of an experimental run in Ainur. Blocks in orange are optional.}\label{paper:olguinmunoz2022airnur:fig:flow}
\end{figure*}

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{publications/2022Ainur/figures/network}
    \caption{Network structure assumed by Ainur.}\label{paper:olguinmunoz2022airnur:fig:network}
\end{figure*}

@@

Conceptually, in Ainur, an experiment is decomposed into a layered structure, consisting of
\begin{enumerate*}[itemjoin={{; }}, itemjoin*={{; and }}]
    \item the distributed application (workload)
    \item computation hosts
    \item communication networks connecting the hosts
\end{enumerate*}.
In \cref{paper:olguinmunoz2022airnur:fig:overview}, the workload is represented by a client-server process pair.
These could correspond, for instance, to the emulation of an inverted pendulum and a matching controller.
However, Ainur makes no assumptions about the nature of the workloads deployed on the framework, and virtually any process or combination of processes can be used.

Hosts correspond to either bare-metal machines or cloud instances on \gls{AWS} \gls{EC2}.
Users are free to combine and interconnect these in any configuration.
Ainur can provision wired networks over ethernet, and supports a number of software-defined wireless communication stacks, currently including Wi-Fi, 4G \gls{LTE}, and 5G.
To realize these various wireless communication protocols, it is assumed that the testbed is equipped with \gls{USRP} \gls{SDR}s and some computation hosts dedicated to signal processing and/or radio management.

@@

    \item[Containers:]
    We leverage Docker containers to great extent for the virtualization and orchestration of workloads, as well as the encapsulation and management of complex network configurations.
    See \cref{paper:olguinmunoz2022airnur:containers} for more details.

@@

\subsection{Containerization}\label{paper:olguinmunoz2022airnur:containers}

@@

Ainur leverages containerization extensively across the framework, in order to
\begin{enumerate*}[itemjoin={{, }}, itemjoin*={{, and }}]
    \item deploy and orchestrate workloads
    \item support a wide spectrum of different physical layer configurations out-of-the-box, as well as allow for easy extension to new ones
    \item support automated collection of logs
\end{enumerate*}.
Workloads in the framework are deployed packaged in Docker containers, orchestrated through \texttt{docker-compose} and Docker Swarm.
We have selected Docker Swarm as our container orchestration layer instead of more advanced toolkits such as Kubernetes due to
\begin{enumerate*}[itemjoin={{; }}, itemjoin*={{; and }}]
    \item it being more lightweight
    \item its comparative simplicity in terms of configuration and setup
    \item the fact that it is included in the default Docker runtime installation
\end{enumerate*}.
This allows the framework to remain mostly agnostic to the nature of the workloads, and thus support a wide ranged of different applications and system architectures.
It also allows for easy deployment to different compute nodes in the network, without having to take into consideration details such as required libraries and versions.

@@

\begin{figure}
    \centering
    \includegraphics[height=25em]{publications/2022Ainur/figures/network-minimal}
    \caption{Minimal testbed setup on which Ainur can be deployed.}\label{paper:olguinmunoz2022airnur:fig:network:minimal}
\end{figure}

@@

A full-featured, and rather complex deployment is depicted in \cref{paper:olguinmunoz2022airnur:fig:network}, including a number of optional testbed features such as offload to the cloud through \gls{VPN} gateways.
However, a minimal a setup can be achieved with a control host, at least two workload hosts, and a couple of managed (smart) switches, as illustrated in \cref{paper:olguinmunoz2022airnur:fig:network:minimal}.
The only requirements for the aforementioned hosts are that a general purpose Linux distribution such as Ubuntu can be deployed on them, and --- in the case of the workload hosts --- the presence of two network interfaces.
Thus, a minimal testbed can be achieved cheaply and quickly by employing small single-board computers such as Raspberry Pi, Jetson Nano, or BeagleBone units.
Finally, the configuration and deployment of an Ainur-compatible testbed can be easily automated with tools such as Ansible, and playbooks for this effect can be made available to the community upon request.

>>>>

\section{The Ainur Framework}\label{paper:olguinmunoz2022ainur:ainur}

\begin{figure}[t]
    \centering
    \includegraphics[width=\textwidth]{publications/2022Ainur/figures/overview}
    \caption{Layered structure of an Ainur experiment}\label{paper:olguinmunoz2022ainur:fig:overview}
\end{figure}

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{publications/2022Ainur/figures/flow2}
    \caption{Lifecycle of an experimental run in Ainur. Blocks in orange are optional.}\label{paper:olguinmunoz2022ainur:fig:flow}
\end{figure*}

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{publications/2022Ainur/figures/network}
    \caption{Network structure assumed by Ainur.}\label{paper:olguinmunoz2022ainur:fig:network}
\end{figure*}

@@

Conceptually, in Ainur, an experiment is decomposed into a layered structure, consisting of
\begin{inlineenum}
    \item the distributed application (workload)
    \item computation hosts
    \item communication networks connecting the hosts
\end{inlineenum}.
In \cref{paper:olguinmunoz2022ainur:fig:overview}, the workload is represented by a client-server process pair.
These could correspond, for instance, to the emulation of an inverted pendulum and a matching controller.
However, Ainur makes no assumptions about the nature of the workloads deployed on the framework, and virtually any process or combination of processes can be used.

Hosts correspond to either bare-metal machines or cloud instances on \gls{AWS} \gls{EC2}.
Users are free to combine and interconnect these in any configuration.
Ainur can provision wired networks over ethernet, and supports a number of software-defined wireless communication stacks, currently including Wi-Fi, 4G \gls{LTE}, and 5G.
To realize these various wireless communication protocols, it is assumed that the testbed is equipped with \gls{USRP} \gls{SDR}'s and some computation hosts dedicated to signal processing and/or radio management.

@@

    \item[Containers:]
    We leverage Docker containers to great extent for the virtualization and orchestration of workloads, as well as the encapsulation and management of complex network configurations.
    See \cref{paper:olguinmunoz2022ainur:containers} for more details.

@@

\subsection{Containerization}\label{paper:olguinmunoz2022ainur:containers}

@@

Ainur leverages containerization extensively across the framework, in order to
\begin{inlineenum}
    \item deploy and orchestrate workloads
    \item support a wide spectrum of different physical layer configurations out-of-the-box, as well as allow for easy extension to new ones
    \item support automated collection of logs
\end{inlineenum}.
Workloads in the framework are deployed packaged in Docker containers, orchestrated through \texttt{docker-compose} and Docker Swarm.
We have selected Docker Swarm as our container orchestration layer instead of more advanced toolkits such as Kubernetes due to
\begin{inlineenum}
    \item it being more lightweight
    \item its comparative simplicity in terms of configuration and setup
    \item the fact that it is included in the default Docker runtime installation
\end{inlineenum}.
This allows the framework to remain mostly agnostic to the nature of the workloads, and thus support a wide ranged of different applications and system architectures.
It also allows for easy deployment to different compute nodes in the network, without having to take into consideration details such as required libraries and versions.

@@

\begin{figure}
    \centering
    \includegraphics[height=25em]{publications/2022Ainur/figures/network-minimal}
    \caption{Minimal testbed setup on which Ainur can be deployed.}\label{paper:olguinmunoz2022ainur:fig:network:minimal}
\end{figure}

@@

A full-featured, and rather complex deployment is depicted in \cref{paper:olguinmunoz2022ainur:fig:network}, including a number of optional testbed features such as offload to the cloud through \gls{VPN} gateways.
However, a minimal a setup can be achieved with a control host, at least two workload hosts, and a couple of managed (smart) switches, as illustrated in \cref{paper:olguinmunoz2022ainur:fig:network:minimal}.
The only requirements for the aforementioned hosts are that a general purpose Linux distribution such as Ubuntu can be deployed on them, and --- in the case of the workload hosts --- the presence of two network interfaces.
Thus, a minimal testbed can be achieved cheaply and quickly by employing small single-board computers such as Raspberry Pi, Jetson Nano, or BeagleBone units.
Finally, the configuration and deployment of an Ainur-compatible testbed can be easily automated with tools such as Ansible, and playbooks for this effect can be made available to the community upon request.

### 2023-02-03 publications/2022Ainur/demo.tex
fixing formatting

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{publications/2022Ainur/figures/demo_hardware}
    \caption[caption]{%
        The possible workload offloading setups and physical layer configurations used in the demo.
        Not pictured, in Phy Option 2:
        \begin{enumerate*}[itemjoin={{; }}, itemjoin*={{; and }}]
            \item \gls{SDR} processing hosts
            \item additional base-station host for the Core Network and \gls{eNodeB}
        \end{enumerate*}.
    }\label{paper:olguinmunoz2022airnur:fig:democonfigs}
\end{figure*}

\begin{figure}
    \centering
    \begin{subfigure}{.45\textwidth}
        \centering
        \includegraphics[width=.9\textwidth]{publications/2022Ainur/figures/demo_workload_1}
        \caption{Workload 1: \acs{MNIST} image classifier}\label{paper:olguinmunoz2022airnur:fig:wkld:mnist}
    \end{subfigure}%
    \hfill%
%    \vspace{1em}
    \begin{subfigure}{.45\textwidth}
        \centering
        \includegraphics[width=.9\textwidth]{publications/2022Ainur/figures/demo_workload_2}
        \caption{Workload 2: inverted pendulum \acs{NCS}}\label{paper:olguinmunoz2022airnur:fig:wkld:ncs}
    \end{subfigure}
    \caption{Demonstration workloads}\label{paper:olguinmunoz2022airnur:fig:wkld}
\end{figure}

\section{Demo}\label{paper:olguinmunoz2022airnur:demo}

@@

\Cref{paper:olguinmunoz2022airnur:fig:democonfigs} illustrates the possible configurations for the testbed.
Workloads are deployed on client hosts, and computation is offloaded either to an edge server (cloudlet), or to \gls{AWS} \gls{EC2} instances on the cloud.
Communication between the client- and server-sides of the workload will occur over one of three possible physical link layer setups:
\begin{description}
    \item[WiFi:] \pgls{SDR} is configured as an IEEE 802.11n access point, and client hosts connect to it using on-board Wi-Fi.
    \item[4G \gls{LTE}:] \pgls{SDR} is configured as an 4G \gls{LTE} base station, and another is configured as an 4G \gls{LTE} \gls{UE}.
    Together these radios act as an \gls{LTE} bridge between the client-side and the server-side of the network, and all client-server traffic is routed through them.
    \item[Ethernet:] connects everything through plain ethernet.
\end{description}

The number of workload instances (and therefore client hosts) deployed in each execution of this demonstration will range from \numrange[]{1}{10}.
Workloads deployed to the cloud will be able to target any of \gls{AWS}s datacenters.

We will employ two different workloads for this demonstration, illustrated in \cref{paper:olguinmunoz2022airnur:fig:wkld}.
The first of these (\cref{paper:olguinmunoz2022airnur:fig:wkld:mnist}) consists of a proof-of-concept web application which implements a simple classifier to identify hand-drawn digits from the widely-used \gls{MNIST} dataset~\cite{deng2012mnist}.
The application consists of a client with a web interface, and \pgls{HTTP} server which responds to requests from the client and performs the actual image recognition.
This workload is intended to showcase in an interactive manner the effects of placing computation at different points of the network, and how Ainur simplifies these deployments.

The second workload (\cref{paper:olguinmunoz2022airnur:fig:wkld:ncs}) corresponds to \pgls{NCS} balancing an inverted pendulum, implemented on a software framework for the emulation of \glspl{NCS} using cloud-native technologies~\cite{olguinmunoz2022cleave}.
It consists of an emulation of the physical inverted pendulum system plant and a software-implemented proportional-differential controller.
These components communicate with each other over the \gls{UDP}.
This workload will be used to showcase the utility of Ainur for automating the execution of batches of experiments potentially including multiple different clients, servers, and physical layers.

@@

This demonstration will be performed on a testbed consisting of
\begin{enumerate*}[itemjoin={{; }}, itemjoin*={{; and finally }}]
    \item \num{10} Raspberry Pi 4 Model B boards, acting as client-side workload hosts
    \item a Raspberry Pi 4 Model B acting as \gls{VPN} gateway for the workload network
    \item a Raspberry Pi 4 Model B acting as \gls{VPN} gateway for the management network, as well as hosting the necessary \gls{NTP} and \gls{DNS} server software
    \item a \texttt{i386} workstation, acting as an edge-side workload host
    \item a configurable number of  \gls{AWS} \gls{EC2} cloud instances acting as cloud servers
    \item a separate \texttt{i386} workstation hosting the Fluent server and on which Ainur is deployed as well.
\end{enumerate*}
These nodes are interconnected using a combination of managed switches, \glspl{SDR}, and \gls{VPN} gateways; please refer to \cref{paper:olguinmunoz2022airnur:fig:network} for an architectural overview of this setup.

@@

\begin{figure}
    \centering
    \includegraphics[height=25em]{publications/2022Ainur/figures/demo_mockup}
    \caption{Screenshot of the user interface of demo workload 1. After deciding on the placement of the compute backend and the type of physical layer connecting client and backend, participants will use this interface to interact with the system.}\label{paper:olguinmunoz2022airnur:fig:mockup}
\end{figure}

Participants will be asked to choose
\begin{enumerate*}[itemjoin={{; }}, itemjoin*={{; and }}]
    \item a location to offload computation to (local (i.e.\ no offloading), edge, or any \gls{AWS} datacenter)
    \item a physical layer for the first hop of the network (Wi-Fi, 4G \gls{LTE}, or ethernet).
\end{enumerate*}
Through the Ainur command line, we will deploy a single client-server pair according to the specifications.
Next, participants will be able to access the client web interface and interact with the server by requesting classification of images from the \gls{MNIST} database.
The server will respond to these, and the assigned labels will be visible on the web interface together with timing statistics for each request, see \cref{paper:olguinmunoz2022airnur:fig:mockup}.

>>>>

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{publications/2022Ainur/figures/demo_hardware}
    \caption[caption]{%
        The possible workload offloading setups and physical layer configurations used in the demo.
        Not pictured, in Phy Option 2:
        \begin{inlineenum}
            \item \gls{SDR} processing hosts
            \item additional base-station host for the Core Network and \gls{eNodeB}
        \end{inlineenum}.
    }\label{paper:olguinmunoz2022ainur:fig:democonfigs}
\end{figure*}

\begin{figure}
    \centering
    \begin{subfigure}{.45\textwidth}
        \centering
        \includegraphics[width=.9\textwidth]{publications/2022Ainur/figures/demo_workload_1}
        \caption{Workload 1: \acs{MNIST} image classifier}\label{paper:olguinmunoz2022ainur:fig:wkld:mnist}
    \end{subfigure}%
    \hfill%
%    \vspace{1em}
    \begin{subfigure}{.45\textwidth}
        \centering
        \includegraphics[width=.9\textwidth]{publications/2022Ainur/figures/demo_workload_2}
        \caption{Workload 2: inverted pendulum \acs{NCS}}\label{paper:olguinmunoz2022ainur:fig:wkld:ncs}
    \end{subfigure}
    \caption{Demonstration workloads}\label{paper:olguinmunoz2022ainur:fig:wkld}
\end{figure}

\section{Demo}\label{paper:olguinmunoz2022ainur:demo}

@@

\Cref{paper:olguinmunoz2022ainur:fig:democonfigs} illustrates the possible configurations for the testbed.
Workloads are deployed on client hosts, and computation is offloaded either to an edge server (cloudlet), or to \gls{AWS} \gls{EC2} instances on the cloud.
Communication between the client- and server-sides of the workload will occur over one of three possible physical link layer setups:
\begin{description}
    \item[Wi-Fi:] \pgls{SDR} is configured as an IEEE 802.11n access point, and client hosts connect to it using on-board Wi-Fi.
    \item[4G \gls{LTE}:] \pgls{SDR} is configured as an 4G \gls{LTE} base station, and another is configured as an 4G \gls{LTE} \gls{UE}.
    Together these radios act as an \gls{LTE} bridge between the client-side and the server-side of the network, and all client-server traffic is routed through them.
    \item[Ethernet:] connects everything through plain ethernet.
\end{description}

The number of workload instances (and therefore client hosts) deployed in each execution of this demonstration will range from \numrange[]{1}{10}.
Workloads deployed to the cloud will be able to target any of \gls{AWS}'s datacenters.

We will employ two different workloads for this demonstration, illustrated in \cref{paper:olguinmunoz2022ainur:fig:wkld}.
The first of these (\cref{paper:olguinmunoz2022ainur:fig:wkld:mnist}) consists of a proof-of-concept web application which implements a simple classifier to identify hand-drawn digits from the widely-used \gls{MNIST} dataset~\cite{deng2012mnist}.
The application consists of a client with a web interface, and \pgls{HTTP} server which responds to requests from the client and performs the actual image recognition.
This workload is intended to showcase in an interactive manner the effects of placing computation at different points of the network, and how Ainur simplifies these deployments.

The second workload (\cref{paper:olguinmunoz2022ainur:fig:wkld:ncs}) corresponds to \pgls{NCS} balancing an inverted pendulum, implemented on a software framework for the emulation of \glspl{NCS} using cloud-native technologies~\cite{olguinmunoz2022cleave}.
It consists of an emulation of the physical inverted pendulum system plant and a software-implemented proportional-differential controller.
These components communicate with each other over the \gls{UDP}.
This workload will be used to showcase the utility of Ainur for automating the execution of batches of experiments potentially including multiple different clients, servers, and physical layers.

@@

This demonstration will be performed on a testbed consisting of
\begin{inlineenum}[itemjoin={{; }}, itemjoin*={{; and finally }}]
    \item \num{10} Raspberry Pi 4 Model B boards, acting as client-side workload hosts
    \item a Raspberry Pi 4 Model B acting as \gls{VPN} gateway for the workload network
    \item a Raspberry Pi 4 Model B acting as \gls{VPN} gateway for the management network, as well as hosting the necessary \gls{NTP} and \gls{DNS} server software
    \item a \texttt{i386} workstation, acting as an edge-side workload host
    \item a configurable number of  \gls{AWS} \gls{EC2} cloud instances acting as cloud servers
    \item a separate \texttt{i386} workstation hosting the Fluent server and on which Ainur is deployed as well.
\end{inlineenum}
These nodes are interconnected using a combination of managed switches, \glspl{SDR}, and \gls{VPN} gateways; please refer to \cref{paper:olguinmunoz2022ainur:fig:network} for an architectural overview of this setup.

@@

\begin{figure}
    \centering
    \includegraphics[height=25em]{publications/2022Ainur/figures/demo_mockup}
    \caption{Screenshot of the user interface of demo workload 1. After deciding on the placement of the compute backend and the type of physical layer connecting client and backend, participants will use this interface to interact with the system.}\label{paper:olguinmunoz2022ainur:fig:mockup}
\end{figure}

Participants will be asked to choose
\begin{inlineenum}
    \item a location to offload computation to (local (i.e.\ no offloading), edge, or any \gls{AWS} datacenter)
    \item a physical layer for the first hop of the network (Wi-Fi, 4G \gls{LTE}, or ethernet).
\end{inlineenum}
Through the Ainur command line, we will deploy a single client-server pair according to the specifications.
Next, participants will be able to access the client web interface and interact with the server by requesting classification of images from the \gls{MNIST} database.
The server will respond to these, and the assigned labels will be visible on the web interface together with timing statistics for each request, see \cref{paper:olguinmunoz2022ainur:fig:mockup}.

### 2023-02-03 publications/2022Ainur/introduction.tex
fixing formatting

Over the last years, several small- to mid-scale testbeds have emerged that leverage a large degree of freedom with respect to hardware and software, for example, the 
\begin{enumerate*}[itemjoin={{, }}, itemjoin*={{, and }}]
    \item \gls{COSMOS}
    \item \gls{POWDER}
    \item \gls{ARA}
    \item Drexel Grid \gls{SDR}
\end{enumerate*} testbeds.
\gls{COSMOS} is a testbed spanning an area of roughly \num{1} square mile (\SI{2.6}{\kilo\meter\squared}) featuring \glspl{SDR}, \si{\milli\meter}-wave equipment, optical fibers, cloud integration, and compute for core network functionality and application data processing~\cite{yu2019cosmos,raychaudhuri2020challenge}.
It contains over \num{200} rooftop, intermediate, and mobile nodes, and is controlled and managed by a central node.
\gls{COSMOS} relies on the \gls{OMF} (originally developed for ORBIT~\cite{ott2005orbit}), and employs the \gls{OEDL}, a domain-specific imperative language based on Ruby, for experiment development and definition.

@@

The rest of this paper is structured as follows.
\cref{paper:olguinmunoz2022airnur:ainur} presents the framework as well as the key concepts and technologies supporting its design and architecture.
This section also discusses briefly the assumptions made about the underlying hardware on which the framework is set to run, and we present an overview of our experimental testbed.
Next, \cref{paper:olguinmunoz2022airnur:demo} describes two demonstration procedures through which we will showcase the flexibility and potential of this tool for the automatic, repeatable, end-to-end experimentation in the context of wireless testbeds.
Finally, in \cref{paper:olguinmunoz2022airnur:conclusion} we summarize our contributions, discuss future directions, and conclude this paper.

>>>>

Over the last years, several small- to mid-scale testbeds have emerged that leverage a large degree of freedom with respect to hardware and software, for example, the 
\begin{inlineenum}
    \item \gls{COSMOS}
    \item \gls{POWDER}
    \item \gls{ARA}
    \item Drexel Grid \gls{SDR}
\end{inlineenum} testbeds.
\gls{COSMOS} is a testbed spanning an area of roughly \num{1} square mile (\SI{2.6}{\kilo\meter\squared}) featuring \glspl{SDR}, \si{\milli\meter}-wave equipment, optical fibers, cloud integration, and compute for core network functionality and application data processing~\cite{yu2019cosmos,raychaudhuri2020challenge}.
It contains over \num{200} rooftop, intermediate, and mobile nodes, and is controlled and managed by a central node.
\gls{COSMOS} relies on the \gls{OMF} (originally developed for ORBIT~\cite{ott2005orbit}), and employs the \gls{OEDL}, a domain-specific imperative language based on Ruby, for experiment development and definition.

@@

The rest of this paper is structured as follows.
\cref{paper:olguinmunoz2022ainur:ainur} presents the framework as well as the key concepts and technologies supporting its design and architecture.
This section also discusses briefly the assumptions made about the underlying hardware on which the framework is set to run, and we present an overview of our experimental testbed.
Next, \cref{paper:olguinmunoz2022ainur:demo} describes two demonstration procedures through which we will showcase the flexibility and potential of this tool for the automatic, repeatable, end-to-end experimentation in the context of wireless testbeds.
Finally, in \cref{paper:olguinmunoz2022ainur:conclusion} we summarize our contributions, discuss future directions, and conclude this paper.

### 2023-02-03 publications/2022Ainur/software_layers/ip_layer.tex
glossary fixes
fixing formatting

\mintinline{python}{LANLayer} handles local network connectivity.
It can be used as a context manager for automatic network teardown, as well as an iterator over the hosts it manages.
This class exposes two main \gls{API} methods, \mintinline{python}{self.add_hosts()} and \mintinline{python}{self.tear_down()}:
\begin{description}[]
    \item[\mintinline{python}{LANLayer.add_hosts(self, layer2)}]
    This method takes a \mintinline{python}{PhysicalLayer} object instance and configures \gls{IP} layer connectivity on the hosts within.
    This includes assigning \gls{IP} addresses to interfaces, configuring default and static routes, as well as authenticating and connecting to the desired access point in the case of WiFi.
    The underlying implementation uses Ansible to write Netplan YAML configuration files on the target hosts and subsequently apply the network configuration.

@@

\begin{itemize}
    \item Two VPNCloud gateways, one for control connections and one for workload data streams, are configured and accessible on the control network.
    \item One of these gateways is additionally connected to the workload data network through a separate interface.
    \item These gateways are both listening on separate \gls{UDP} ports on a single public IP address.
    In other words, both gateways access the public internet through a single \gls{NAT} gateway, on which appropriate \gls{UDP} port forwarding rules are configured to allow \gls{VPN} traffic back into the control network.
    \item The VPNCloud connections configured on the the gateways share the same pre-shared key.
    \item The VPNCloud links belong to mutually \emph{disjoint} \acs{IP}v4 address ranges.
    % \item The gateways are assigned the first host address in the available address space of their respective \gls{VPN} links.
    % For example, if the gateway \gls{VPN} link address space is \texttt{10.0.0.0/24}, the gateway has address \texttt{10.0.0.1} on the \gls{VPN} link.
    \item Devices on the control network have been preconfigured with appropriate \gls{IP} routes to reach the control gateway and its associated \gls{VPN} link.
    \item Finally, cloud instances intended to be used with this class have VPNCloud installed on them.
\end{itemize}

>>>>

\mintinline{python}{LANLayer} handles local network connectivity.
It can be used as a context manager for automatic network teardown, as well as an iterator over the hosts it manages.
This class exposes two main \gls{API} methods, \mintinline{python}{self.add_hosts()} and \mintinline{python}{self.tear_down()}:
\begin{description}[]
    \item[\mintinline{python}{LANLayer.add_hosts(self, layer2)}]
    This method takes a \mintinline{python}{PhysicalLayer} object instance and configures \gls{IP} layer connectivity on the hosts within.
    This includes assigning \gls{IP} addresses to interfaces, configuring default and static routes, as well as authenticating and connecting to the desired access point in the case of Wi-Fi.
    The underlying implementation uses Ansible to write Netplan YAML configuration files on the target hosts and subsequently apply the network configuration.

@@

\begin{itemize}
    \item Two VPNCloud gateways, one for control connections and one for workload data streams, are configured and accessible on the control network.
    \item One of these gateways is additionally connected to the workload data network through a separate interface.
    \item These gateways are both listening on separate \gls{UDP} ports on a single public IP address.
    In other words, both gateways access the public internet through a single \gls{NAT} gateway, on which appropriate \gls{UDP} port forwarding rules are configured to allow \gls{VPN} traffic back into the control network.
    \item The VPNCloud connections configured on the the gateways share the same pre-shared key.
    \item The VPNCloud links belong to mutually \emph{disjoint} \gls{IP}v4 address ranges.
    % \item The gateways are assigned the first host address in the available address space of their respective \gls{VPN} links.
    % For example, if the gateway \gls{VPN} link address space is \texttt{10.0.0.0/24}, the gateway has address \texttt{10.0.0.1} on the \gls{VPN} link.
    \item Devices on the control network have been preconfigured with appropriate \gls{IP} routes to reach the control gateway and its associated \gls{VPN} link.
    \item Finally, cloud instances intended to be used with this class have VPNCloud installed on them.
\end{itemize}

### 2023-02-03 publications/2022Ainur/software_layers/physical.tex
fixing formatting

This layer creates and deploy the underlying physical connections of the data network and is implemented in the \mintinline{python}{PhysicalLayer} class. 
Types of connections handled by this layer could be ethernet and WiFi links, or even an entire software-defined 4G \gls{LTE} network.

@@

Currently, WiFi and \gls{LTE} implementations are available.
User-defined implementation for other wireless connections are possible and encouraged.

@@

\begin{description}[style=nextline]
    \item[SDR-based WiFi Network]
    The \mintinline{python}{SDRWifi} class handles the configuration and deployment of a software-defined IEEE 802.11 network using the Mangocomm software package and its supported software-defined radios \footnote{See https://mangocomm.com/802.11-mac-phy/}.
    It lunches a Docker container on the same machine to communicate with the \glspl{SDR} over the control network.
    This communication is for the configuration and monitoring the WiFi nodes.
    In a WiFi network, a radio is either \emph{station} or \emph{access point}.
    To initialize a \mintinline{python}{SDRWifi} class, the user is required to provide two lists containing the access points and station nodes instances as described below.

    \mintinline{python}{APSoftwareDefinedRadio} class represents a WiFi access point.
    It advertises a network with a specific SSID on a radio channel and waits for the incoming connections.

    \mintinline{python}{StationSoftwareDefinedRadio} class is used for the station WiFi nodes.
    It searches for an WiFi \gls{SSID} among the networks that are being advertised and attempts for establishing the connection.

    When a WiFi network is up and an station is connected, the data network interfaces of the radios behave as if an ethernet cable is connecting them.
    Thus, the client hosts and the server hosts can be on the same ip layer subnets.
    \begin{figure}[t]
        \centering
        \includegraphics[width=0.8\linewidth]{figures/wifi.png}
        \caption{An example of WiFi network topology in data network and its corresponding \glspl{VLAN}}
        \label{fig:wifi}
    \end{figure}

@@

    \gls{LTE} \gls{EPC} consists of HSS, MME, SPGW-U, and SPGW-C services which can be run on the same host.
    For each \gls{eNodeB} or \gls{UE}, a software-defined radio and a host for signal processing must be specified additionally.
    The resulting topology of an \gls{LTE} network in Ainur is depicted in Figure \ref{fig:lte}.
    Unlike WiFi wireless connection, due to the fact that \gls{LTE} creates an internal ip network and that introduces routing complexities, the server host and the client host cannot be on the same subnet in the ip layer.

>>>>

This layer creates and deploy the underlying physical connections of the data network and is implemented in the \mintinline{python}{PhysicalLayer} class. 
Types of connections handled by this layer could be ethernet and Wi-Fi links, or even an entire software-defined 4G \gls{LTE} network.

@@

Currently, Wi-Fi and \gls{LTE} implementations are available.
User-defined implementation for other wireless connections are possible and encouraged.

@@

\begin{description}[style=nextline]
    \item[SDR-based Wi-Fi Network]
    The \mintinline{python}{SDRWifi} class handles the configuration and deployment of a software-defined IEEE 802.11 network using the Mangocomm software package and its supported software-defined radios \footnote{See https://mangocomm.com/802.11-mac-phy/}.
    It lunches a Docker container on the same machine to communicate with the \glspl{SDR} over the control network.
    This communication is for the configuration and monitoring the Wi-Fi nodes.
    In a Wi-Fi network, a radio is either \emph{station} or \emph{access point}.
    To initialize a \mintinline{python}{SDRWifi} class, the user is required to provide two lists containing the access points and station nodes instances as described below.

    \mintinline{python}{APSoftwareDefinedRadio} class represents a Wi-Fi access point.
    It advertises a network with a specific SSID on a radio channel and waits for the incoming connections.

    \mintinline{python}{StationSoftwareDefinedRadio} class is used for the station Wi-Fi nodes.
    It searches for an Wi-Fi \gls{SSID} among the networks that are being advertised and attempts for establishing the connection.

    When a Wi-Fi network is up and an station is connected, the data network interfaces of the radios behave as if an ethernet cable is connecting them.
    Thus, the client hosts and the server hosts can be on the same ip layer subnets.
    \begin{figure}[t]
        \centering
        \includegraphics[width=0.8\linewidth]{figures/wifi.png}
        \caption{An example of Wi-Fi network topology in data network and its corresponding \glspl{VLAN}}
        \label{fig:wifi}
    \end{figure}

@@

    \gls{LTE} \gls{EPC} consists of HSS, MME, SPGW-U, and SPGW-C services which can be run on the same host.
    For each \gls{eNodeB} or \gls{UE}, a software-defined radio and a host for signal processing must be specified additionally.
    The resulting topology of an \gls{LTE} network in Ainur is depicted in Figure \ref{fig:lte}.
    Unlike Wi-Fi wireless connection, due to the fact that \gls{LTE} creates an internal ip network and that introduces routing complexities, the server host and the client host cannot be on the same subnet in the ip layer.

### 2023-02-03 publications/2022CLEAVE/conclusion.tex
fixing formatting

We validate the utility of this tool through an example use case approximating a proposed edge deployment of inverted pendula control loops co-located with video analytics services.
We argue such a use case represents a realistic scenario and appropriate benchmark for the tool, since
\begin{enumerate*}[itemjoin={{; }}, itemjoin*={{; and }}]
    \item the inverted pendulum plant is ubiquitous in \gls{NCS} research
    \item similar setups exist in real-world industrial use
    \item video analytics has long been proposed as a ``killer app'' for edge computing.
\end{enumerate*}
Our results showcase the ability of the framework to extract relevant metrics relating to the stability of the control system, as well as on the performance of the underlying network link.
We believe \gls{CLEAVE} represents an important step towards enabling inexpensive and low-complexity scalable research for real-world deployment of edge-bound \glspl{NCS}.

>>>>

We validate the utility of this tool through an example use case approximating a proposed edge deployment of inverted pendula control loops co-located with video analytics services.
We argue such a use case represents a realistic scenario and appropriate benchmark for the tool, since
\begin{inlineenum}
    \item the inverted pendulum plant is ubiquitous in \gls{NCS} research
    \item similar setups exist in real-world industrial use
    \item video analytics has long been proposed as a ``killer app'' for edge computing.
\end{inlineenum}
Our results showcase the ability of the framework to extract relevant metrics relating to the stability of the control system, as well as on the performance of the underlying network link.
We believe \gls{CLEAVE} represents an important step towards enabling inexpensive and low-complexity scalable research for real-world deployment of edge-bound \glspl{NCS}.
